---
title: "Assess Normality"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Assess Normality}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"#,
  # fig.path = "man/figures/assess-normality-",
  # out.width = "100%"
)

options(knitr.kable.NA = '')
hcolor <- "#363B74"
tcolor <- "#EEEEEE"
```

```{r setup}
library(MetaPipe)
```

`MetaPipe` assesses the normality of variables (traits) by performing a 
Shapiro-Wilk test on the raw data (see [Load Raw Data](load-raw-data.html) and 
[Replace Missing Data](replace-missing-data.html)). Based on whether or not the
data approximates a normal distribution, an array of transformations will be 
computed, and the normality assessed one more time.

```{r assess-normality-example-data-before, echo = FALSE, dpi = 150, fig.align = 'center'}
set.seed(123)
test_data <- rnorm(500)
MetaPipe::generate_hist(exp(1)^test_data, NULL, xlab = latex2exp::TeX("e^{data}"), save = FALSE)
```

The diagram below shows the tree of transformations that can be performed, the
user can specify the transformation values passing a vector to `transf_vals`; by 
default, `[2, e, 3, 4, 5, 6, 7, 8, 9, 10]`.
```{r assess-normality-diagram, echo = FALSE}
DiagrammeR::grViz(system.file("figures/assess-normality-graph.gv", package = "MetaPipe"), width = "100%")
```