---
title: "Replace Missing Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replace Missing Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/replace-missing-data-",
  out.width = "100%"
)
```

```{r setup}
library(MetaPipe)
```

`MetaPipe` can handle with missing data in a couple of ways:

1. Drop traits which exceed an NA proportion threshold (50% by default)
2. Replace NAs by half of the minimum value within each trait 

## Function call
```{r replace-missing-data-fx-call, eval = FALSE}
replace_missing(raw_data = raw_data, 
                excluded_columns = c(2, 3, ..., M), 
                # Optional
                out_prefix = "metapipe", 
                prop_na = 0.5, 
                replace_na = FALSE)
```
where `raw_data` is a data frame containing the raw data, as described in [Load Raw Data](load-raw-data.html) and `excluded_columns` is a vector containing the indices of the properties, e.g. `c(2, 3, ..., M)`. The other arguments are optional, `out_prefix` is the prefix for output files, `prop_na` is the proportion of NA values (used to drop traits), and `replace_na` is a logical flag to indicate whether or not NAs should be replace by half of the minimum value.

## Drop traits
By default the pipeline will drop traits that exceed an NA proportion threshold, 
this can be fine tuned by the user with the parameter `prop_na`. It is important 
to keep in mind the side effects of excluding variables from the QTL mapping, 
like making wrong conclusions regarding the most significant QTLs.

```{r drop-traits-example}
set.seed(123)
example_data <- data.frame(ID = 1:5,
                           P1 = c("one", "two", "three", "four", "five"), 
                           T1 = rnorm(5), 
                           T2 = rnorm(5),
                           T3 = c(NA, rnorm(4)),                     #  20 % NAs
                           T4 = c(NA, 1.2, -0.5, NA, 0.87),          #  40 % NAs
                           T5 = NA)                                  # 100 % NAs

workdir <- tempdir()
out_prefix <- file.path(workdir, "metapipe")

# Default parameters: NA proportion = 50%
replace_missing(example_data, c(2), out_prefix = out_prefix)

# NA proportion = 30%
replace_missing(example_data, c(2), prop_na = 0.3, out_prefix = out_prefix)
```

## Replace NAs
Alternatively, the user can indicate whether or not NA values should be replaced
by finding the minimum value for each trait and dividing it by two. This result 
can be achieved by passing the parameter `replace_na = TRUE`. Users should be 
cautious when using this approach, as the replacement of data points can have
side effects that might lead to the wrong conclusions.

```{r replace-na-example}
set.seed(123)
example_data <- data.frame(ID = 1:5,
                           P1 = c("one", "two", "three", "four", "five"), 
                           T1 = rnorm(5), 
                           T2 = rnorm(5),
                           T3 = c(NA, rnorm(4)),                     #  20 % NAs
                           T4 = c(NA, 1.2, -0.5, NA, 0.87),          #  40 % NAs
                           T5 = NA)                                  # 100 % NAs

workdir <- tempdir()
out_prefix <- file.path(workdir, "metapipe")

# Default parameters: NA proportion = 50%
replace_missing(example_data, c(2), replace_na = TRUE, out_prefix = out_prefix)

# NA proportion = 30%
replace_missing(example_data, 
                c(2), 
                prop_na = 0.3, 
                replace_na = TRUE, 
                out_prefix = out_prefix)
```

From the last example can be seen that `prop_na` and `replace_na` are mutually
exclusive, and `replace_na` has precedence over `prop_na`.

Next, see [Assess Normality](assess-normality.html). 

```{r clean-knitting-output, echo = FALSE, warning = FALSE, message = FALSE}
filenames <- c(list.files(workdir, "example_*"))
output <- lapply(filenames, file.remove)
```