---
title: "Replace Missing Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replace Missing Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(MetaPipe)
```

### Create toy dataset
```{r loading-raw-data}
# Toy dataset
set.seed(123)
example_data <- data.frame(ID = c(1,2,3,4,5),
                           P1 = c("one", "two", "three", "four", "five"), 
                           F1 = rnorm(5), 
                           F2 = rnorm(5))
```

### Function call
```{r replace-missing-data-fx-call, eval = FALSE}
replace_missing(raw_data = example_data, 
                excluded_columns = c(2), 
                # Optional
                out_prefix = "metapipe", prop_na = 0.5, replace_na = FALSE)
```
where `raw_data` is a data frame containing the raw data, as described in [Load Raw Data](load-raw-data) and `excluded_columns` is a vector containing the indices of the properties, e.g. `c(2, 3, ..., M)`. The other arguments are optional, `out_prefix` is the prefix for output files, `prop_na` is the proportion of NA values (used to drop features), and `replace_na` is a logical flag to indicate whether or not NAs should be replace by half of the minimum value.

```{r replace-missing-data-example}
# Inserting missing values manually
example_data$F1[2:3] <- NA
example_data$F2[4] <- NA

# No changes expected
replace_missing(example_data, c(2))

# Traits with 25% of NA should be dropped
replace_missing(example_data, c(2), prop_na =  0.25)

# NAs should be replaced by half of the minimum value
replace_missing(example_data, c(2), replace_na =  TRUE)
```